\textit{Supporting Information for}: Geometric models reveal behavioral and neural signatures of how naturalistic experiences are transformed into episodic memories

### 2020-08-20 paper/supporting_information.tex
minor edits to supplement

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
\usepackage{placeins}
%\doublespacing

@@

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over episode sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-subject episode-recall trajectory correlations and per-subject hand-annotated memory performance ratings.}
\label{fig:paramsearch}
\end{figure}

@@

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the structure of the episode topic proportions, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode topic trajectory (created using all hand-annotated features from the 998 hand-annotated scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with episode trajectories created using all but one type of feature.  We created temporal correlation matrices for each trajectory (using the topic proportions matrices) and correlated the proximal diagonals from the upper triangles of each impoverished trajectory with the original feature-complete trajectory (see \textit{Methods} for details on how we isolated proximal temporal correlations).  Observing a lower correlation between an impoverished trajectory (holding out a particular feature) and the feature-complete trajectory would suggest that the given feature played a more prominent role in shaping the temporal structure of the feature-complete trajectory.  We found that hand-annotated narrative details provided the most structure to the feature-complete trajectory, whereas the name of the character(s) in focus for each shot provided the least structure (Fig.~\ref{fig:feature-importance}A).

@@

We also carried out an analysis of which annotated features tended to shape aspects of the episode topic trajectory's structure that were preserved in participants' recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode topic trajectory, and correlated the proximal diagonals from its upper triangle with those of the timepoint-by-timepoint correlation matrices for each participant's recall topic trajectory (resampled using linear interpolation to have the same number of timepoints as the episode trajectory).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall trajectories (when a given feature was held out) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized type of feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between scenes, the camera angle, the names of the various characters on screen, and which character was speaking tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall topic vector correlation than including them; Fig.~\ref{fig:feature-importance}B).

Next, we wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed episode topic trajectories from annotations of each individual feature, in turn, and then compared the proximal diagonals from the upper triangles of the temporal correlation matrices of all pairs of features.  This provided additional confirmation that the shape of the full trajectory (including all types of features) was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic trajectories (Fig.~\ref{fig:feature-importance}C).
%\FloatBarrier

\subsection*{Creating a low-dimensional embedding space}
Figures~7 and 8C in the main text display two-dimensional projections of the 100-dimensional topic trajectories for the episode (7A, 8C), average recall (7B), and each individual's recall (7C, 8C).  We created these embeddings using the Uniform Manifold Approximation and Projection algorithm \citep[UMAP;][]{McInEtal18} called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  An advantage of the UMAP algorithm over comparable manifold learning techniques (such as $t$-SNE) is that UMAP explicitly attempts to preserve the global structure of the data \citep{McInEtal18,BechEtal19} by constructing a space where distance on the manifold is standard Euclidean distance, with respect to the global coordinate system.  This was important in our use case, as we wanted to visualize both the evolving structure of the episode and the spatial relationships between presented and recalled content.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{placeins}

@@

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over episode sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-subject episode-recall trajectory correlations and per-subject hand-generated memory performance scores.}
\label{fig:paramsearch}
\end{figure}

@@

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the shape of the episode topic trajectory, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode topic trajectory (created using all hand-annotated features from the 998 manually identified scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with episode trajectories created using all but one type of feature.  For each impoverished episode trajectory, we computed the timepoint-by-timepoint correlation matrix, and correlated the proximal diagonals from the upper right triangle with those from the temporal correlation matrix of the feature-complete trajectory (for details on how we isolated proximal temporal correlations, see \textit{Methods}).  Observing a lower correlation between an impoverished trajectory (with a particular feature removed) and the feature-complete trajectory would suggest that the held-out feature contributed more prominently to the full episode trajectory's temporal structure.  We found that hand-annotated narrative details played the greatest roll in determining the shape of the episode, whereas the name of the character(s) in focus for each shot contributed least (Fig.~\ref{fig:feature-importance}A).

@@

Next, we sought to determine which annotated features contributed aspects of the episode trajectory's shape that were preserved in participants' later recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode topic trajectory, and correlated the proximal diagonals from its upper triangle with those from the timepoint-by-timepoint correlation matrices for each participant's recall topic trajectory (stretched via linear interpolation to have the same number of timepoints as the episode trajectory).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall trajectories (constructed in the absence of a given feature) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between scenes, the camera angle, the names of the various characters on screen, and the location in which the scene took place tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall topic vector correlation than including them; Fig.~\ref{fig:feature-importance}B).

We also wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed topic trajectories from the annotations for each individual feature, in turn, and (using the same technique as in the above analyses) compared the proximal temporal correlation structure of each single-feature trajectory to each other, as well as to that of the full episode.  This provided additional confirmation that the shape of the full episode trajectory was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic trajectories (Fig.~\ref{fig:feature-importance}C).
%\FloatBarrier

\subsection*{Creating a low-dimensional embedding space}
Figures~7 and 8C in the main text display two-dimensional projections of the 100-dimensional topic trajectories for the episode (7A, 8C), average recall (7B), and each individual's recall (7C, 8C).  We created these embeddings using the Uniform Manifold Approximation and Projection algorithm \citep[UMAP;][]{McInEtal18} called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  An advantage of the UMAP algorithm over comparable manifold learning techniques (e.g., $t$-SNE) is that UMAP explicitly attempts to preserve the global structure of the data \citep{McInEtal18,BechEtal19} by constructing a space where distance on the manifold is standard Euclidean distance, with respect to the global coordinate system.  This was important in our use case, as we wanted to visualize both the evolving structure of the episode and the spatial relationships between presented and recalled content.
